---
title: "cachematrix"
author: "kassem saleh"
date: "March 18, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown



### This Rmd file contains two functions to illustrate the use of a cached object
### to avoid recomuptation of the inverse of a matrix if the matrix did not change
### the two functions are: makeCacheMatrix and cacheSolve



### makeCacheMatrix creates a special matrix object that can cache its inverse
### this function returns a list of functions as follows:
###    setMatrix() to set the values inside the matrix
###    getMatrix() to get the values stored inside the matrix
###    getInverse() to get the values of the inverse of the matrix
###    cacheInverse() to get the cached values of the inverse of the matrix

```{r}
makeCacheMatrix <- function(x = matrix()) {
    
    # initially, set cache to NULL since no inverse was computed first
    cache <- NULL
    
    # store new values in matrix
    setMatrix <- function(newMatrixValues) {
      x <<- newMatrixValues
      
      # set cache to NULL since new values are stored in matrix
      cache <<- NULL
    }
    
    # returns the stored matrix
    getMatrix <- function() {
      x
    }
    
    # returns the cached matrix
    getInverse <- function() {
      cache
    }
    
    
    # cache the given argument 
    cacheInverse <- function(solve) {
      cache <<- solve
    }
    
    # return a list. Each named element of the list is a function
    list(setMatrix = setMatrix, 
         getMatrix = getMatrix, 
         cacheInverse = cacheInverse, 
         getInverse = getInverse)
  }


## cacheSolve computes the inverse of the special matrix object returned by makeCacheMatrix. 
## if the inverse matrix has been computed previously and the matrix has not changed, then
## this function retrieves the inverse from the cache
## we assume that the matrix object is invertible

cacheSolve <- function(x, ...) {

    # get the inverse from cache in case it was computed before
    inverse <- x$getInverse()
    
    # if previously cached return it from cache
    if(!is.null(inverse)) {
      message("getting cached data")
      
      return(inverse)
    }
    
    # if did not return the cached value (i.e., null inverse, not computed before)
    # Get the matrix, and compute its inverse and store it in cache
    data <- x$getMatrix()
    inverse <- solve(data)
    x$cacheInverse(inverse)
    
    # return the inverse
    inverse
  }
  
### test if the caching works
```


### first make a 3x3 invertible matrix
```{r}


a <- makeCacheMatrix()


a$setMatrix(matrix(c(7,2,1,0,3,-1,-3,4,-2), nrow = 3, ncol = 3) )
```
### show that matrix

```{r}
a$getMatrix()
```


### invert that matrix -- show a message if cache was used -- should not because first time
```{r}
cacheSolve(a)
```


### now invert again the same matrix  -- message should indicate cache was used
```{r}
cacheSolve(a)
```

### now use another matrix and show it
```{r}
a$setMatrix( matrix(c(7,0,-3,2,3,4,1,-1,-2), nrow = 3, ncol = 3) )

a$getMatrix()
```

### solve again -- cache not used 
```{r}
cacheSolve(a)

```
### now try the same with 2x2 invertible matrix

```{r}

a <- makeCacheMatrix()

a$setMatrix( matrix(c(1,5,10,25), nrow = 2, ncol = 2))

a$getMatrix()

cacheSolve(a)

cacheSolve(a)

a$setMatrix (matrix(c(1,5,10,20), nrow = 2, ncol = 2))

cacheSolve(a)
```

